---
title: "simpleRegress"
author: "Mairani G"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

For model length~neck, adjested $r^2$ =0.75   

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age


```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

For model length~age, adjusted $r^2$ =0.5073, this is less than the adjusted $r^2$ for neck so this model is not as good as the neck model.  

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```
For model length~chest, adjusted $r^2$ =0.7869 , this is a little bit better than the adjusted $r^2$ for AGE so this model is slightly better than the AGE model. It is also better than the NECK model that = to 0.75.   


```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

For model length~HEADLEN, adjusted $r^2$ =0.8434  , this is a little bit better than the adjusted $r^2$ for CHEST so this model is slightly better than the CHEST model(which was= to 0.7869).   

```{r}
headwthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthModel)
summary.lm(headwthModel)
```
For model length~HEADWTH, adjusted $r^2$ =0.5321   , this is worse than the adjusted $r^2$ for HEADLEN so this model is worse than the chest model.  

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

For model length~WEIGHT, the adjusted $r^2$ =0.7423, this is better than the adjusted $r^2$ for HEADWTH so this model is better than the HEADWTH model. 

It is not the best model however, HEADLEN is still better as its adjusted $r^2$= 0.8434.  

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

For model length~SEX, the adjusted $r^2$ =-0.008373 , this is a lot worse than the adjusted $r^2$ for WEIGHT so this model is worse than the WEIGHT model.  

## Conclusion

Based on the prediction of LENGTH, the best model was HEADLEN. Its value of Adjusted-R squared was better than the rest as it was bigger. Its $r^2$ was equal to 0.8434, bigger than the second highest which was CHEST at 0.7869.  